---
title: "Prediction"
author: "Imogen Schwandner -2671952S"
date: '2022-08-17'
output: html_document
---

Predictions using the habitat suitability model developed in "RSF.Rmd" across the study area to derive a habitat suitability raster which can then be converted into a resistance surface for use in connectivity analysis in Circuitscape.

1. Load libraries

```{r}
library(raster)
library(rgeos)
library(rgdal)
library(sp)

```

2. Load model and covariate raster layers

```{r}
# habitat suitability model
# (will already exist in environment if "4_RSF.Rmd" is worked through first)
load("mod.RData")

# covariate rasters
load("WBS_covariates.RData")

# data frame with occurrences and pseudo-absences as well as extracted covariates
# (will already exist in environment if "4_RSF.Rmd" is worked through first)
load("RSFdf.RData")
```

3. Creating raster stack to make predictions on

3.1. Creating NDVI and DNDVI layer

3.1.1. NDVI

NDVI not treated as dynamic in prediction, but stationary, based on average NDVI of the region between 2000 and 2022

```{r}
#read mean NDVI raster file
# edit filepaths as needed
NDVI <- raster("./Smithsonian_data/Data/Data_Layers/32736/MEAN_MOD13Q1_250m_16_days_NDVI_2022-06-08.tif")

# re-project raster to project crs (epsg:32736)
NDVI <- projectRaster(NDVI, crs = crs)

# crop to extent of study area (anthropogenic risk layer)
NDVI <- crop(NDVI, anthro)

# re-sample raster to resolution of remaining layers using bilinear interpolation (250m down to ~51 m)
NDVI<-resample(NDVI, anthro, method="bilinear")
```

3.1.2. DNDVI

DNDVI also not treated as dynamic. Here I use the difference between two mean 16-day NDVI images from the beginning to the late middle of March 2011. The middle of March is the beginning of the long wet season, when green-up (positive change in NDVI) is expected, but before the wettest month (April) with the most extreme changes in green-up. Data collection in 2010 did not start until the end of May, when senescence can already be observed, hence the choice of 2011 data (which still constitute pre-fencing circumstances).

```{r}
#load NDVI layers for the beginning and end of march 2011 (16 day mean at 250m resolution)
# edit file paths as needed
M1 <- raster("./Smithsonian_data/Data/Data_Layers/32736/MOD13Q1.A2011065.250m_16_days_NDVI.tif") # beginning of March '11
M2 <- raster("./Smithsonian_data/Data/Data_Layers/32736/MOD13Q1.A2011081.250m_16_days_NDVI.tif") # middle/end of March '11

# Take the difference between the end of March and beginning of March 16 day mean (DNDVI)
DNDVI <- M2-M1

#re-project
DNDVI <- projectRaster(DNDVI, crs = crs)

# crop to study area extent
DNDVI <- crop(DNDVI, anthro)

# re-sample to desired resolution
DNDVI<-resample(DNDVI, anthro, method="bilinear")

```

3.2. Stack and scale all covariate layers

```{r}
# create raster stack
stack <- stack(NDVI, DNDVI, TWI, D_wood, D_river, D_roadT1, D_roadT5, anthro)

#scale
stack <- scale(stack)

#change names of stack to match the model
names(stack) <- c("indvi", "dndvi", "TWI", "Dwood", "Driver", "Drdt1", "Drdt5", "anthroRisk")
```

4.   Make predictions

4.1. Create grid to make predictions on

```{r}
# extent for grid
e <- extent(stack)

# coerce to a SpatialPolygons object
p <- as(e, 'SpatialPolygons')

# set resolution
RES <- res(stack)

#create grid
grid <- makegrid(p, cellsize = RES)

# coerce to SPDF
grid <- SpatialPointsDataFrame(grid,
                               data = data.frame(ID=1:nrow(grid)))
#set crs
grid@proj4string <- stack@crs
```

4.2. Extract grid values from ratser stack

```{r}
#extract covariates
temp <- raster::extract(stack, grid)

# add data
grid@data<-as.data.frame(temp)
# dummy variable for AID
grid@data$AID <- "a"
```

4.3. Prediction

4.3.1. Add Predictions

```{r}
# add predictions
grid$pred <- predict(mod, re.form = NA, # ignores random effect when making predictions
                          newdata = grid@data,
                          type = "response")
```

4.3.2. Convert greid with predictions to raster

```{r}
# turn into raster
pred_rast <- rasterFromXYZ(xyz = data.frame(x = coordinates(grid)[,1],
                                            y = coordinates(grid)[,2],
                                            z = grid$pred),
                           res = c(RES,RES),
                           crs = grid@proj4string)
# plot prediction raster
# plot(pred_rast)
```

4.3.3. Save to file

```{r}
# save predicted habitat suitability as GeoTiff file
writeRaster(pred_rast, file = "pred_rast.tif")
```


