---
title: "9_Validation"
author: "Imogen Schwandner -2671952S"
date: "2022-08-24"
output: html_document
---

1. Load libraries

```{r message=FALSE, warning=FALSE}
library(raster)
library(glmmTMB)
library(lme4)
library(rgdal)
library(sf)
```

2. Load wildebeest movement data sets (2010-2013 and 2017-2021) and pseudo-absence points for both

# 2017-2021 movement data are not yet publicly available and therefore no longer included in this repository

```{r}
# load movement data sets
load("WB_and_Covariates_data.RData")

# remove covariates not needed here
rm(anthro, D_river, D_roadT1, D_roadT5, D_wood, TWI)

## Load restoration area of interest extent for fine scale validation
#load restoration extend
Rextent <- readOGR("restoration_extent.shp")


# load 2010-2013 pseudo-absences (created in "2_Pseudo_Absences-1013.Rmd")
load("Points.RData")

# rename
Points1013 <- Points

# isolate AID column needed for GLMMs
P1013AID <- Points1013[,2]

# reduce Points to x and y coordinates for raster::extract
Points1013 <- Points1013[,3:4]

# load 2017-2021 pseudo-absences (created in "8_Pseudo_Absences -1721validation.Rmd")
load("Points1721.RData")

# rename
Points1721 <- Points

##Crop pseudo-absences (Points1721) for fine scale validation in restoration AOI
# Convert the data frame to an sf object
Points1721_sf <- st_as_sf(Points1721, coords = c("x", "y"), crs = st_crs(Rextent))
# Crop the points using the bounding box
Points1721R <- st_crop(Points1721_sf, st_bbox(Rextent))
# Convert the sf object to a data frame without geometry
Points1721R <- cbind(st_drop_geometry(Points1721R), st_coordinates(Points1721R))

# isolate AID column needed for GLMMs (full regional scale)
P1721AID <- Points1721[,2]

# reduce Points to x and y coordinates for raster::extract (full regional scale)
Points1721 <- Points1721[,3:4]

# isolate AID column needed for GLMMs (restoration fine scale)
P1721AIDR <- Points1721R[,2]

# reduce Points to x and y coordinates for raster::extract (restoration fine scale)
Points1721R <- Points1721R[,3:4]

rm(Points)

##Crop WBH data to restoration extent for fine scale validation

#ensure CRS of data and extent match
crs(Rextent) <- crs(WBH)

#create restoration extent WBH locations by cropping WBH by restoration extent
WBHR <- crop(WBH, extent(Rextent))

```


3. Read in models to validate

3.1. Habitat Suitability Surface
```{r}
# read in predicted habitat suitability raster for study area
suitability <-  raster("pred_rast.tif")
```

3.2. Pre-Fencing Connectivity Model
```{r}
#read in pre-fence connectivity (cumulative current) map
prefence <- raster("PREFENCE_MCP_N_MM_CS_cum_curmap.asc")
```

3.3. Difference map (Fenced - Pre-Fencing connectivity levels)

```{r}
# read in fenced connectivity levels
fence <- raster("FENCE_polygon_MCP_N_MM_CS_cum_curmap.asc")

# subtract historic levels form current levels (fence - pre-fence) to obtain difference map where losses in connectivity show up as negatives and gains as positives
diffconn <- fence - prefence
```


4.     Validation

4.1.   Extract variables for validation models

4.1.1. Habitat Suitability

GOAL: Validate predicted habitat suitability using WBH (2017-2021) movement data occurrences and pseudo-absences

Regional scale (entire Mara)
```{r}
# extract habitat suitability values for occurrences
pres <- raster::extract(suitability, WBH)

# create data frame with suitability value, AID and Presence/Absence column (1 = presence)
presAID <-  data.frame("Suitability" = pres, "AID" = WBH@data[,1], "PA" = 1)

# extract habitat suitability values for pseudo-absences
abs <-  raster::extract(suitability, Points1721)

# create data frame with suitability value, AID and Presence/Absence column (0 = absence)
absAID <- data.frame("Suitability" = abs, "AID" = P1721AID, "PA" = 0)

# combine presence and absence data frames for modelling
WBHPoints1721 <- rbind(presAID, absAID)

# ensure PA is a factor for modelling
WBHPoints1721$PA <- as.factor(WBHPoints1721$PA)

```

Fine scale (restoration area of interest)
```{r}
# extract habitat suitability values for occurrences
pres <- raster::extract(suitability, WBHR)

# create data frame with suitability value, AID and Presence/Absence column (1 = presence)
presAIDR <-  data.frame("Suitability" = pres, "AID" = WBHR@data[,1], "PA" = 1)

# extract habitat suitability values for pseudo-absences
abs <-  raster::extract(suitability, Points1721R)

# create data frame with suitability value, AID and Presence/Absence column (0 = absence)
absAIDR <- data.frame("Suitability" = abs, "AID" = P1721AIDR, "PA" = 0)

# combine presence and absence data frames for modelling
WBHPoints1721R <- rbind(presAIDR, absAIDR)

# ensure PA is a factor for modelling
WBHPoints1721R$PA <- as.factor(WBHPoints1721R$PA)

```

4.1.2. Pre-Fence Connectivity

GOAL: Validate pre-fencing connectivity model using WBH (2017-2021) movement data occurrences and pseudo-absences

Regional scale (entire Mara)
```{r}
# extract connectivity value (cumulative current) for occurrences
pres2 <- raster::extract(prefence, WBH)

# extract connectivity value (cumulative current) for pseudo-absences
abs2 <-  raster::extract(prefence, Points1721)

# combine values
C <- c(pres2,abs2)

# add to WBHPoints1721 data frame
WBHPoints1721$Current <- C

```

Fine scale (restoration extent)
```{r}
# extract connectivity value (cumulative current) for occurrences
pres2 <- raster::extract(prefence, WBHR)

# extract connectivity value (cumulative current) for pseudo-absences
abs2 <-  raster::extract(prefence, Points1721R)

# combine values
C <- c(pres2,abs2)

# add to WBHPoints1721 data frame
WBHPoints1721R$Current <- C

```

4.1.3. Connectivity Difference Map

GOAL: Validate connectivity change difference map using WBH (2017-2021) movement data occurrences and pseudo-absences

Regional scale (entire Mara)
```{r}
# extract connectivity loss/gain at occurrences
pres3 <- raster::extract(diffconn, WBH)

# extract connectivity loss/gain at pseudo-absences
abs3 <- raster::extract(diffconn, Points1721)

# combine values
CC <- c(pres3, abs3)


# add connectivity change results to WBHPoints1721R data frame
WBHPoints1721$CChange <- CC

```

Fine scale (restoration area of interest)
```{r}
# extract connectivity loss/gain at occurrences
pres4 <- raster::extract(diffconn, WBHR)

# extract connectivity loss/gain at pseudo-absences
abs4 <- raster::extract(diffconn, Points1721R)

# combine values
CC <- c(pres4,abs4)

# add to WBHPoints1721 data frame
WBHPoints1721R$CC <- CC

```

4.2. Remove NA's

Some pseudo-absences (less than 0.003 % of observations) were created outside of the study area extend due to large buffer sizes in some individuals. These create NA's in variable extraction which interfere with modelling and hence need to be removed.

```{r}
# removes entire rows with NA's created in absences due to some points coming from buffers outside of the range of the study area
WBHPoints1721 <- WBHPoints1721[complete.cases(WBHPoints1721),]
```

Less than 1 % of data points removed.

Remove NA's on fine scale restoration validation
```{r}
# removes entire rows with NA's created in absences due to some points coming from buffers outside of the range of the fine scale validation
WBHPoints1721R <- WBHPoints1721R[complete.cases(WBHPoints1721R),]
```

About 1% of data removed.

4.3. Modelling

######################################################
######### VALIDATING PRE-FENCE SUITABILITY ###########
######################################################

Regional scale (entire Mara)
```{r}
### GLMM ###
# model the effect of presence or absence on suitability accounting for variation in animal ID (AID) with a logit link from a beta distribution
modTMBS <- glmmTMB(Suitability ~ PA + (1|AID), WBHPoints1721, family=list(family="beta",link="logit"))
summary(modTMBS)
```

RESULT: 
Predicted habitat suitability is significantly higher (0.55 ± 0.0057 times higher) at occurrences than at pseudo-absences of peri-fencing movement data.

Fine scale (restoration area of interest)
```{r}
### GLMM ###
# model the effect of presence or absence on suitability accounting for variation in animal ID (AID) with a logit link from a beta distribution
modTMBSR <- glmmTMB(Suitability ~ PA + (1|AID), WBHPoints1721R, family=list(family="beta",link="logit"))
summary(modTMBSR)

```
```{r}
#fixing the error
#disproves the error
# sum(is.na(WBHPoints1721R$Suitability))
# sum(is.nan(WBHPoints1721R$Suitability))
# sum(is.infinite(WBHPoints1721R$Suitability)) 

modTMBSR1 <- glmmTMB(Suitability ~ PA + (1|AID), 
                    data = WBHPoints1721R, 
                    family = list(family = "beta", link = "logit"),
                    start = list(beta = c(0, 0), b = rep(0, length(unique(WBHPoints1721R$AID)))))
```


```{r}
# tried several things to fix this
# fitting the model without the random effect works
modTMBSRsimple <- glmmTMB(Suitability ~ PA, WBHPoints1721R, family=list(family="beta",link="logit"))
summary(modTMBSRsimple)
# logit transforming suitability to get a "normal" slightly right-skewed distribution to then fit an lmer model works
WBHPoints1721R$logit_Suitability <- log(WBHPoints1721R$Suitability / (1 - WBHPoints1721R$Suitability))
hist(WBHPoints1721R$logit_Suitability)


modlmSR <- lmer(logit_Suitability ~ PA + (1|AID), WBHPoints1721R)
summary(modlmSR)

```

######################################################
######### VALIDATING PRE-FENCE CURRENT ###############
######################################################

Regional scale (entire Mara)
```{r}
### LMM ###
# model the effect of presence or absence on cumulative current (connectivity) - current is log transformed to normalise - using a linear mixed effects model and a random effect for animal ID
modlmC <- lmer(log(Current) ~ PA + (1|AID), WBHPoints1721)
summary(modlmC)
```

RESULT:
Current is 0.47 ± 0.0048 times higher at presences than pseudo-absences of peri-fencing movement data. 

Fine scale (restoration area of interest)
```{r}
### LMM ###
# model the effect of presence or absence on cumulative current (connectivity) - current is log transformed to normalise - using a linear mixed effects model and a random effect for animal ID
modlmCR <- lmer(log(Current) ~ PA + (1|AID), WBHPoints1721R)
summary(modlmCR)
```
RESULT:
Validation data wildebeest moved in areas where pre-fencing connectivity was significantly higher (0.077772 ± 0.007567) than areas available to them. This underlines that the validation data wildebeest still closely tracked movements of the 2010-2013 historic pre-fencing wildebeest over the entire Mara region. However, at the finer scale of the restoration area of interest, we observe a much weaker coefficient underlining that in this area, where many historic pathways were already blocked by fencing in 2017-2021, validation data wildebeest already moved in areas that were less important for connectivity historically. 


######################################################
######### VALIDATING CONNECTIVITY DIFFERENCE #########
######################################################

Regional scale (entire Mara)
```{r}
### LMM ###
# model the effect of presence or absence on current/connectivity change accounting for variation in animal ID (AID) from a normal distribution
modlmCC <- lmer(CChange ~ PA + (1|AID), WBHPoints1721)
summary(modlmCC)
```


RESULT:
Peri-fencing wildebeest moved in areas of significantly higher (0.000239 ± 0.00001 times higher) connectivity gain (rather than loss) compared to pre-fencing wildebeest. Hence the model predicted the changes in movement well.

Fine scale (restoration area of interest)
```{r}
### LMM ###
# model the effect of presence or absence on current/connectivity change accounting for variation in animal ID (AID) from a normal distribution
modlmCCR <- lmer(CC ~ PA + (1|AID), WBHPoints1721R)
summary(modlmCCR)
```
  -0.0007179000
SE 0.0000276300

RESULT:
Validation data Wildebeest (2017-2021) within the restoration area of interest move predominantly in areas of significant connectivity loss (-0.0007179 ± 0.0000276 lower at presences than pseudo-absences) than expected by chance.


######################################################
#########  VALIDATING  FENCE  RESISTANCE     #########
######################################################

```{r}
#load resistance surface with fences and fenced in polygons at resistance 100
resistance <- raster("C:/Users/Owner/Desktop/MScQMBCE/DISS/Circuitscape/occupanczfenceplr.asc")
```

Regional scale (entire Mara)
```{r}
# extract connectivity loss/gain at occurrences
pres5 <- raster::extract(resistance, WBH)

# extract connectivity loss/gain at pseudo-absences
abs5 <- raster::extract(resistance, Points1721)

# combine values
R <- c(pres5, abs5)


# add connectivity change results to WBHPoints1721R data frame
WBHPoints1721$Resistance <- R

# get proportion of presences that fall on fences (resistance = 100)

```

Fine scale (restoration area of interest)
```{r}
# extract connectivity loss/gain at occurrences
pres4 <- raster::extract(diffconn, WBHR)

# extract connectivity loss/gain at pseudo-absences
abs4 <- raster::extract(diffconn, Points1721R)

# combine values
CC <- c(pres4,abs4)

# add to WBHPoints1721 data frame
WBHPoints1721R$CC <- CC

```