---
title: "9_Validation"
author: "Imogen Schwandner -2671952S"
date: "2022-08-24"
output: html_document
---

1. Load libraries

```{r}
library(raster)
```

2. Load wildebeest movement data sets (2010-2013 and 2017-2021) and pseudo-absence points for both

# 2017-2021 movement data are not yet publicly available and thereofore not availbale on this repository

```{r}
# load movement data sets
load("WB_and_Covariates_data.RData")
# remove covariates not needed here
rm(anthro, anthro1, D_river, D_roadT1, D_roadT5, D_wood, TWI, WB)

# load 2010-2013 pseudo-absences (created in "2_Pseudo_Absences-1013.Rmd")
load("Points.RData")

# rename and keep only x and y column
Points1013 <- Points[,3:4]

# load 2017-2021 pseudo-absences (created in "8_Pseudo_Absences -1721validation.Rmd")
load("Points1721.RData")

# rename and keep only x and y column
Points1721 <- Points[,3:4]
```

3. Read in models to validate

3.1. Habitat Suitability Surface
```{r}
# read in predicted habitat suitabilty raster for study area
suitability <-  raster("pred_rast.tif")
```

3.2. Pre-Fencing Connectivity Model
```{r}
#read in pre-fence connectivity (cumulative current) map
prefence <- raster("PREFENCE_MCP_N_MM_CS_cum_curmap.asc")
```

3.3. Difference map (Fenced - Pre-Fencing connectivity levels)

```{r}
# read in fenced connectivity levels
fence <- raster("FENCE_polygon_MCP_N_MM_CS_cum_curmap.asc")

# subtract historic levels form current levels (fence - pre-fence) to obtain difference map where losses in connectivity show up as negatives and gains as positives
diffconn <- fence - prefence
```


4.    Validation
4.1.  Habitat Suitability

######################################################
######### VALIDATING PRE-FENCE SUITABILITY ###########
######################################################

Using WBH (2017-2021) movement data occurrences and pseudo-absences

```{r}
# extract habitat suitability values for occurrences
pres <- raster::extract(suitability, WBH)

# extract habitat suitability values for pseudo-absences
abs <-  raster::extract(suitability, Points1721)

# t-test to compare group means
t.test(pres, abs)
```

Suitability is significantly higher at occurrences than at pseudo-absences.


4.2. Pre-Fence Connectivity

######################################################
######### VALIDATING PRE-FENCE CURRENT ###############
######################################################

```{r}
# extract connectivity (cumulative current) for occurrences
pres2 <- raster::extract(prefence, WBH)

# extract connectivity (cumulative current) for pseudo-absences
abs2 <-  raster::extract(prefence, Points1721)

# t-test to compare group means
t.test(pres2, abs2)
```

4.3. Connectivity Difference Map

######################################################
######### VALIDATING CONNECTIVITY DIFFERENCE #########
######################################################

```{r}
# extract connectivity loss/gain at occurrences
pres3 <- raster::extract(diffconn, WBH)

# extract connectivity loss/gain at pseudo-absences
abs3 <- raster::extract(diffconn, Points1721)

# t-test to compare group means
t.test(pres3, abs3)
```


4.4. Compare habitat suitability between movement data sets

```{r}
# extract habitat suitability at 2017-2021 occurrences (peri-fencing)
hsWBH <- raster::extract(suitability, WBH)

# extract habitat suitability at 2010-2013 occurrences (pre-fencing)
hsWBS <- raster::extract(suitability, WBS)

# t-test to compare group means
t.test(hsWBH, hsWBS)
```

