---
title: "9_Validation"
author: "Imogen Schwandner -2671952S"
date: "2022-08-24"
output: html_document
---

1. Load libraries

```{r message=FALSE, warning=FALSE}
library(raster)
library(glmmTMB)
library(lme4)
```

2. Load wildebeest movement data sets (2010-2013 and 2017-2021) and pseudo-absence points for both

# 2017-2021 movement data are not yet publicly available and therefore no longer included in this repository

```{r}
# load movement data sets
load("WBS_covariates.RData")


# remove covariates not needed here
rm(anthro, D_river, D_roadT1, D_roadT5, D_wood, TWI)


# load 2010-2013 pseudo-absences (created in "2_Pseudo_Absences-1013.Rmd")
load("Points.RData")

# rename
Points1013 <- Points

# isolate AID column needed for GLMMs
P1013AID <- Points1013[,2]

# reduce Points to x and y coordinates for raster::extract
Points1013 <- Points1013[,3:4]

# load 2017-2021 pseudo-absences (created in "8_Pseudo_Absences -1721validation.Rmd")
load("Points1721.RData")
# rename
Points1721 <- Points

# isolate AID column needed for GLMMs
P1721AID <- Points1721[,2]

# reduce Points to x and y coordinates for raster::extract
Points1721 <- Points1721[,3:4]

rm(Points)
```

3. Read in models to validate

3.1. Habitat Suitability Surface
```{r}
# read in predicted habitat suitability raster for study area
suitability <-  raster("pred_rast.tif")
```

3.2. Pre-Fencing Connectivity Model
```{r}
#read in pre-fence connectivity (cumulative current) map
prefence <- raster("PREFENCE_MCP_N_MM_CS_cum_curmap.asc")
```

3.3. Difference map (Fenced - Pre-Fencing connectivity levels)

```{r}
# read in fenced connectivity levels
fence <- raster("FENCE_polygon_MCP_N_MM_CS_cum_curmap.asc")

# subtract historic levels form current levels (fence - pre-fence) to obtain difference map where losses in connectivity show up as negatives and gains as positives
diffconn <- fence - prefence
```


4.     Validation

4.1.   Extract variables for validation models

4.1.1. Habitat Suitability

GOAL: Validate predicted habitat suitability using WBH (2017-2021) movement data occurrences and pseudo-absences


```{r}
# extract habitat suitability values for occurrences
pres <- raster::extract(suitability, WBH)

# create data frame with suitability value, AID and Presence/Absence column (1 = presence)
presAID <-  data.frame("Suitability" = pres, "AID" = WBH@data[,1], "PA" = 1)

# extract habitat suitability values for pseudo-absences
abs <-  raster::extract(suitability, Points1721)

# create data frame with suitability value, AID and Presence/Absence column (0 = absence)
absAID <- data.frame("Suitability" = abs, "AID" = P1721AID, "PA" = 0)

# combine presence and absence data frames for modelling
WBHPoints1721 <- rbind(presAID, absAID)

# ensure PA is a factor for modelling
WBHPoints1721$PA <- as.factor(WBHPoints1721$PA)

```


4.1.2. Pre-Fence Connectivity

GOAL: Validate pre-fencing connectivity model using WBH (2017-2021) movement data occurrences and pseudo-absences


```{r}
# extract connectivity value (cumulative current) for occurrences
pres2 <- raster::extract(prefence, WBH)

# extract connectivity value (cumulative current) for pseudo-absences
abs2 <-  raster::extract(prefence, Points1721)

# combine values
C <- c(pres2,abs2)

# add to WBHPoints1721 data frame
WBHPoints1721$Current <- C

```



4.1.3. Connectivity Difference Map

GOAL: Validate connectivity change difference map using WBH (2017-2021) movement data occurrences and pseudo-absences


```{r}
# extract connectivity loss/gain at occurrences
pres3 <- raster::extract(diffconn, WBH)

# extract connectivity loss/gain at pseudo-absences
abs3 <- raster::extract(diffconn, Points1721)

# combine values
CC <- c(pres3, abs3)

# add to WBHPoints1721 data frame
WBHPoints1721$CChange <- CC
```

4.2. Remove NA's

Some pseudo-absences (less than 0.003 % of observations) were created outside of the study area extend due to large buffer sizes in some individuals. These create NA's in variable extraction which interfere with modelling and hence need to be removed.

```{r}
# removes entire rows with NA's created in absences due to some points coming from buffers outside of the range of the study area
WBHPoints1721 <- WBHPoints1721[complete.cases(WBHPoints1721),]
```

Less than 1 % of data points removed.

4.3. Modelling

######################################################
######### VALIDATING PRE-FENCE SUITABILITY ###########
######################################################

```{r}
### GLMM ###
# model the effect of presence or absence on suitability accounting for variation in animal ID (AID) with a logit link from a beta distribution
modTMBS <- glmmTMB(Suitability ~ PA + (1|AID), WBHPoints1721, family=list(family="beta",link="logit"))
```

RESULT: 
Predicted habitat suitability is significantly higher (0.55 ± 0.0057 times higher) at occurrences than at pseudo-absences of peri-fencing movement data.


######################################################
######### VALIDATING PRE-FENCE CURRENT ###############
######################################################

```{r}
### LMM ###
# model the effect of presence or absence on cumulative current (connectivity) - current is log transformed to normalise - using a linear mixed effects model and a random effect for animal ID
modlmC <- lmer(log(Current) ~ PA + (1|AID), WBHPoints1721)
summary(modlmC)
```

RESULT:
Current is 0.47 ± 0.0048 times higher at presences than pseudo-absences of peri-fencing movement data. 


######################################################
######### VALIDATING CONNECTIVITY DIFFERENCE #########
######################################################

```{r}
### LMM ###
# model the effect of presence or absence on current/connectivity change accounting for variation in animal ID (AID) from a normal distribution
modlmCC <- lmer(CChange ~ PA + (1|AID), WBHPoints1721)
```


RESULT:
Peri-fencing wildebeest moved in areas of significantly higher (0.000239 ± 0.00001 times higher) connectivity gain (rather than loss) compared to pre-fencing wildebeest. Hence the model predicted the changes in movement well.


